---
title: "Preprocessing Directives"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{preprocessor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(rslates)
```


# Introduction

Slate blueprints are just regular R source code files decorated with *slates* preprocessing directives. These directives define the inputs that will appear on the slate, the outputs and their source code, datasets imported, and so on. 

The *Slates* preprocessor is responsible to convert **decorated code** to regular R source code that can be parsed and executed by the R interpreter by substituting in the source code the user-defined values of the inputs. 

# Quick Start

Here we show a summary of all supported preprocessor directives and their usual syntax. Detailed descriptions of each directive are presented in the sections following.

```{r, eval = FALSE}
# Single-hash comment. Will appear in R source and may 
# contain preprocessor directives.
 
## Double-hash comment. Will be suppressed from the R source
## code but may contain preprocessor directives.

### Triple-hash comment. Will be suppressed from R source 
### and preprocessor directives will be ignored.

##
## Input
##
## 'type' is one of character, expression, logical, numeric, choices, numeric2, numeric4
## 
$@input name

$@input name(type, default)
$@input name(type, default, description, parent, ...)

##
## Group
##
$@group id(title, description, layout, parent)
$@group title, description, layout, parent

##
## Page
##
$@page id(title, description, layout, parent)
$@page title, description, layout, parent

##
## Outputs
##
## 'type' is one of print, plot, table, html, markdown.
##
$@output type, title, ...

$end-output

##
## Variable Substitution
##
## 'options' are combinations of x, n, N, d, D, q, Q, lxx.
## 'parameters' follow the same syntax as $@input, but parenthesis are mandatory.
##
## Substitute a single variable
$$var
$$options:var
$$options:var(parameters)

## Substitute multiple variable and separate values with commas
${var1, var2}
${options:: options1:var1, options2:var2}
${options:: options1:var1(parameters1), options2:var2(parameters2)}

##
## Imports
##
$@import name(type)
$@import name, type

## Execute code on import
$@begin-import name(type)
$@begin-import name, type

$@end-import

##
## Exports
##
$@export varname

## Execute code on export
$@begin-export varname
$@end-export

##
## Code Visibility
##
$@invisible
$@visible

##
## Value Setter
##
set_input(name, value)
```


# Example

We start with a complete example showing off most features of the preprocessing engine. 

```{r, eval = FALSE}
$@title "Scatterplot"
$@tags "example", "rgraphics", "scatterplot"
$@preamble FALSE
 
## $@input x(expression, "1:10", description = "x coordinates for the plot; or structure defining both x and y coordinates.")
## $@input y(expression, description = "y coordinates for the plot.")
## 
## $@input type(choices, "p",
##              choices = list("Points"="p", "Lines"="l", "Both"="b", "Empty Points"="c",
##                             "Overplotted"="o", "Stairs (Floor)"="s", "Stairs (Ceiling)"="S"),
##              description = "desired type of plot.")
## 
## $@input main(character, description = "main title for the plot.")
## $@input xlab(character, description = "label for the x-axis.")
## $@input ylab(character, description = "label for the y-axis.")
## 
## $@input xlim(numeric2, c(NA, NA), description = "x limits (x1, x2) for the plot.")
## $@input ylim(numeric2, c(NA, NA), description = "y limits (x1, x2) for the plot.")

$@output plot(Scatterplot)

# set the margins for the plot
par(mar=c(${mar_bottom, mar_left, mar_top, mar_right}))

# do the plot
plot(${dn:: x:x, y=y(check.valid = FALSE), type=type, main=main, xlab=xlab, ylab=ylab})

$@end-output

```



# Comments

Comments come in three flavors (`#`, `##` and `###`). These affect what comments will be *displayed*, *suppressed* and/ or *interpreted* by the preprocessor.

## Single Hash (`#`)

A regular R comment. This comment will show up in the final source that is provided to the user. Any preprocessor directives included in the comment are interpreted by the preprocessor and removed from the source. This type of comment is mostly used to provide the user with an explanation of what the resulting source code is doing.

## Double Hash (`##`)

Will suppressed from user code, but preprocessing directives are interpreted. This type of comment is mostly used to assure that included preprocessor directives do not appear in the generated source code.

## Triple Hash (`###`)

These are used to comment the blueprint source code itself, as they will be suppressed from the user code and ignored by the preprocessor.

# Inputs, Groups and Pages

Inputs, groups and pages are defined through the `$@input`, `$@group` and `$@page` directives. After evaluation, these directives are scrubbed from the user code, and so can be placed outside of comments. However, placing them in double-hash comments may facilitate code-formatting by avoiding unnecessary empty lines in the user code.

## Inputs

> `$@input name(type, default, ...)`

Inputs are defined using the syntax `$@input name(type, default, ...)` where `name` is the variable name associated with the input, `type` is the input type, and `default` is the default value of the input. 

Input types include `character`, `numeric`, `logical`, `choices`, `expression`, `numeric2` and `numeric4`. 

### Usage

In the simplest case, an input may be defined with just a name. In this case it will assume the default `character` type. The default value will depend on the input type.

```
## defines an input 'x' of type 'character' (the default input type).
## $@input x
```

The *input type* and *default value* can be specified with or without parameter names.

```
## defines a 'numeric' input with the name 'num_samples' and default value 10.
## $@input num_samples(numeric, 10)
## $@input num_samples(type = numeric, default = 10)
```

Additional parameters must be specified by name. Common user-interface options include `description`, that determines the tooltip text for the input and `pretty.name` that, if provided, will be used in the user-interface instead of the variable name.

```
## $@input x(expression, expression(1:10), description="x coordinates for the plot.", pretty.name = "x-coordinates")
```
*Note:* for convenience it is not required to quote the input type parameter, although one may do so if desired. `$@input x(character)` is equivalent to `$@input x("character")`.


## Groups

> `$@group name(layout, ...)`

# Outputs and Datasets

> `$@output name(type, ...)`
> ... 
> `$@end-output`



> `$@dataset name(type, ...)`
> ... 
> `$@end-dataset`

# Variable Substitution

> `${options:varname}`

> `${default_options:: options1:varname1, options2:varname2, ...}`

# Conditional Output

> `$@if expression` 
> ... 
> `$@end-if`

# Imports

Import types: 

* *file*: a file on the local computer.
* *rdata*: a Rdata file to be loaded.

```
$@import file filename [end]

$@end-import
```




